---
title: "P8105_hw1_ab6169"
author: "Amrutha Banda"
output: github_document
date: "2025-09-20"
---

## Problem 1 

I first started by installing the necessary packages. 
```{r include=FALSE}
library(tidyverse)
library(moderndive)
library(ggplot2)
```

After this, I loaded the early_january_weather dataset.
```{r}
data("early_january_weather")
```

Here are the variables in the dataset. 
```{r}
names(early_january_weather)
```

I looked at the size of the dataset by seeing how many rows and columns there are.
```{r}
nrow(early_january_weather)
ncol(early_january_weather)
```

Then, I found the mean temperature 
```{r}
mean(early_january_weather$temp, na.rm = TRUE)
```

This is a scatterplot of temperature versus time, with color representing humidity.
```{r}
ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) +
  geom_point(alpha = 0.6, size = 1.2) +
  labs(
    title = "Temperature vs Time (Early January)",
    x = "Time (hourly)",
    y = "Temperature (°F)",
    color = "Humidity"
  )

ggsave("scatter_plot_tempvstime.pdf", height=4, width=6)
```


In Problem 1, I looked at the `early_january_weather` dataset from the `moderndive` package. The important variables in this dataset include `origin`, `year`, `month`, `day`, `hour`, `time_hour`, `temp`, `dewp`, `humid`, `wind_dir`, `wind_speed`, `wind_gust`, `precip`, `pressure`, and `visib`. This dataset has `r nrow(early_january_weather)` rows and `r ncol(early_january_weather)` columns. The mean temperature in this dataset is `r mean(early_january_weather$temp, na.rm = TRUE)` F. This value represents the average air temperature recorded across all observations in early January. To better understand temperature vs time, I created a scatterplot. This scatterplot shows that the temperatures have increased over the two week period. I also notice that humidity varies across time but does not show a strong consistent trend in this plot. Humidity is lower in the beginning of the month of January. As the month progresses, humidity increases a bit. I see lighter blue data points clustered near the Janurary 14th time. 


## Problem 2 

I created a dataframe. 

```{r}
weather_df <- 
  tibble(
    norm_sample = rnorm(10),
    vec_logical = norm_sample > 0,
    vec_char    = c("a","b","c","d","e","f","g","h","i","j"),
    vec_factor  = factor(c("low","medium","high","low","medium",
                           "high","low","medium","high","low"))
  )
```

Then, I took the mean of the variables `norm_sample` and `vec_logical`. Taking the mean of `vec_char` and `vec_factor` returned an error because characters and factors aren't numerical and cannot be averaged. 
```{r}
mean(pull(weather_df, norm_sample))
mean(pull(weather_df, vec_logical))

```


When I write a code chunk that applies the `as.numeric`, I see that  `as.numeric(weather_df$vec_factor)` and `as.numeric(weather_df$vec_logical)`, return numbers. Applying `as.numeric` to the `vec_char` variable returns NA's with a warning about coercion. This is because letters have no numeric meaning. This explains why you cannot take their mean. Yes, this does explain what happened when I tried to take the mean. The mean works for numeric and logical variables but not for character an factor variables because they don't have a numeric meaning.
```{r echo=TRUE, results= 'hide' }
as.numeric(weather_df$vec_logical)
as.numeric(weather_df$vec_char)
as.numeric(weather_df$vec_factor)
```


